A Heuristic Approach towards Drawings of Graphs with High Crossing Resolution

### 2018-05-27 paper/paper.tex
A bit more on the related work section.

\begin{abstract}
The \emph{crossing resolution} of a non-planar drawing of a graph is the minimum angle formed by any pair of crossing edges. Recent experiments have shown that the larger the crossing resolution is, the easier it is to read and interpret a drawing of a graph. However, maximizing the crossing resolution turns out to be an NP-hard problem in general and only heuristic algorithms are known that are mainly based on appropriately adjusting spring embedding algorithms.

@@

From a practical point of view, we are only aware of two methods that aim at drawings with high crossing resolution; both of them are adjustments of spring embedding algorithms~\cite{DBLP:journals/congnum/Eades84}. The first one is due to Huang, Eades, Hong and Lin~\cite{DBLP:journals/vlc/HuangEHL13}, while the second one is due to Argyriou, Bekos and Symvonis~\cite{DBLP:journals/cj/ArgyriouBS13}. Common in both algorithms is that they apply appropriate forces on the endvertices of every pair of crossing edges. Each of the two algorithms uses a different way to compute (the direction and the magnitude of) the forces, but the underlying idea in both algorithms is the same: the smaller the corresponding crossing angles are, the larger are the magnitudes of the forces applied at their endvertices.

@@

In general, randomization is a technique that has not been deeply examined in Graph Drawing, as it seems to be difficult to compute the expected quality of the produced drawings; a notable exception is the randomized approach by Goldschmidt and Takvorian~\cite{DBLP:journals/networks/GoldschmidtT94} for computing large planar subgraphs. Since we could not provide any theoretical guarantee on the expected quality of the produced drawings as well (mainly due to the nature of the problem itself), we decided to follow a more practical approach. To this end, we implemented our algorithm and the ones presented in~\cite{DBLP:journals/vlc/HuangEHL13} and~\cite{DBLP:journals/cj/ArgyriouBS13}, and we experimentally compared the crossing resolutions of the drawings produced by them for standard benchmark graphs (such as the Rome and the North graphs) that are widely used in Graph Drawing for comparing different drawing algorithms. Our experimental evaluation indicates that our new method significantly outperforms the ones that are based on adjustments of spring embedding algorithms~\cite{DBLP:journals/vlc/HuangEHL13,DBLP:journals/cj/ArgyriouBS13} in terms of crossing resolution, but this comes at the cost of slightly worse running time, especially for large and dense graphs.

@@

An immediate observation emerging from the above literature overview is that the focus of the study of the crossing resolution maximization problem has been primarily on theoretical aspects of the problem. The absence of approaches for computing drawings with high crossing resolution in practice is rather obvious and it can be justified, up to a certain point, by the fact that the problem turns to be NP-hard, when asking for drawings with optimal crossing resolution. Exceptions are the works of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou at al.~\cite{DBLP:journals/cj/ArgyriouBS13}, which are both based on the spring-embedding technique~\cite{DBLP:journals/congnum/Eades84}. According to this technique, a graph is modelled as a physical system with forces acting on it, and a (good) drawing is obtained by an equilibrium of the system; for an introduction and a detailed discussion of several different variants refer to~\cite{DBLP:books/ph/BattistaETT99}.

>>>>

\begin{abstract}
The \emph{crossing resolution} of a non-planar drawing of a graph is the minimum angle formed by any pair of crossing edges. Recent experiments have shown that the larger the crossing resolution is, the easier it is to read and interpret a drawing of a graph. However, maximizing the crossing resolution turns out to be an NP-hard problem in general and only heuristic algorithms are known that are mainly based on appropriately adjusting force-directed algorithms.

@@

From a practical point of view, we are only aware of two methods that aim at drawings with high crossing resolution; both of them are adjustments of force-directed algorithms~\cite{DBLP:journals/congnum/Eades84}. The first one is due to Huang, Eades, Hong and Lin~\cite{DBLP:journals/vlc/HuangEHL13}, while the second one is due to Argyriou, Bekos and Symvonis~\cite{DBLP:journals/cj/ArgyriouBS13}. Common in both algorithms is that they apply appropriate forces on the endvertices of every pair of crossing edges. Each of the two algorithms uses a different way to compute (the direction and the magnitude of) the forces, but the underlying idea in both algorithms is the same: the smaller the corresponding crossing angles are, the larger are the magnitudes of the forces applied at their endvertices.

@@

In general, randomization is a technique that has not been deeply examined in Graph Drawing, as it seems to be difficult to compute the expected quality of the produced drawings; a notable exception is the randomized approach by Goldschmidt and Takvorian~\cite{DBLP:journals/networks/GoldschmidtT94} for computing large planar subgraphs. Since we could not provide any theoretical guarantee on the expected quality of the produced drawings as well (mainly due to the nature of the problem itself), we decided to follow a more practical approach. To this end, we implemented our algorithm and the ones presented in~\cite{DBLP:journals/vlc/HuangEHL13} and~\cite{DBLP:journals/cj/ArgyriouBS13}, and we experimentally compared the crossing resolutions of the drawings produced by them for standard benchmark graphs (such as the Rome and the North graphs) that are widely used in Graph Drawing for comparing different drawing algorithms. Our experimental evaluation indicates that our new method significantly outperforms the ones that are based on adjustments of force-directed algorithms~\cite{DBLP:journals/vlc/HuangEHL13,DBLP:journals/cj/ArgyriouBS13} in terms of crossing resolution, but this comes at the cost of slightly worse running time, especially for large and dense graphs.

@@

An immediate observation emerging from the above literature overview is that the focus of the study of the crossing resolution maximization problem has been primarily on theoretical aspects of the problem. Most of the approaches that could be useful in practise are based on adjustments of force-directed techniques~\cite{DBLP:journals/congnum/Eades84}, according to which a graph is modelled as a physical system with forces acting on it, and a (good) drawing is obtained by an equilibrium of the system; for an introduction and a discussion of several variants refer to~\cite{DBLP:books/ph/BattistaETT99}. %Actually, the absence of approaches for computing drawings with high crossing resolution in practice is rather obvious and it can be justified, up to a certain point, by the fact that the problem turns to be NP-hard, when asking for drawings with optimal crossing resolution.

More concretely, from an applicative point of view, Didimo et al.~\cite{DBLP:conf/apvis/DidimoLR10} describe a system, called \emph{COWA}, to support conceptual web site traffic analysis, whose algorithmic core is a force-directed heuristic to compute simultaneous embeddings of two non-planar graphs with high crossing resolution. 
%
In a follow up work, Didimo et al.~\cite{DBLP:conf/gd/DidimoLR10} describe heuristics, designed within the topology-driven force-directed framework, to achieve good trade-offs in terms of number of edge crossings, crossing resolution, and geodesic edge tendency. 
%
However, the obtained drawings are not straight-line. For straight-line drawings, Nguyen et al.~\cite{DBLP:conf/gd/NguyenEHH10} suggest a quadratic-programming based approach to increase the crossing angles of circular drawings. 
%
Of more general scope are the already mentioned works of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou at al.~\cite{DBLP:journals/cj/ArgyriouBS13}, which are both based on the force-directed technique.
